---
title: 'Class 9: Unsupervised Learning II'
author: "Mary Garcia"
date: "April 30, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Exploratory data analysis
##Preparing the data
```{r}
#Assign file to an object
fna.data <- "WisconsinCancer.csv"
#Open object with read.csv
wisc.df <- read.csv(fna.data)
```

```{r}
#View() opens in new tab; head() opens in same file
View(wisc.df)
head(wisc.df)
```

```{r}
#
wisc.data <- as.matrix(wisc.df[3:32])
row.names(wisc.data) <- wisc.df$id
head(wisc.data)
```

Store answers as numeric (1 being malignant)
```{r}
diagnosis <- as.numeric(wisc.df$diagnosis == "M")
```


Q1
```{r}
nrow(wisc.df)
```

Q2
```{r}
colnames(wisc.data)
grep("_mean", colnames(wisc.data))
length(grep("_mean", colnames(wisc.data)))
```


Q3
```{r}

```



## Principal Component Analysis

```{r}
round(colMeans(wisc.data), 1)
```

```{r}
round(apply(wisc.data,2,sd), 1)
```

```{r}
wisc.pr <- prcomp(wisc.data,scale = TRUE)
summary(wisc.pr)
```


```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis+1)
```






Q4
```{r}

```

Q5
```{r}

```

Q6
```{r}

```

##Intrpret PCA results

```{r}
biplot(wisc.pr)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis+1, xlab = "PC1", ylab = "PC2")
```



```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,3], col=diagnosis+1, xlab = "PC1", ylab = "PC3")
```





##Variance explained

```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
```

```{r}
pve <- pr.var/(sum(pr.var))*100
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 100), type = "o")
```


```{r}
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```


```{r}
par(mfcol=c(1,2))
pve <- pr.var/(sum(pr.var))*100
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 100), type = "o")

plot(cumsum(pve) , xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 100), type = "o")
```



## Hierarchial clustering of case data

```{r}
data.scaled <- scale(wisc.data)
```

```{r}
data.dist <- dist(data.scaled)
```

```{r}
wisc.hclust <- hclust(data.dist, method = "complete")
```

```{r}
plot(wisc.hclust)
abline(h = 19, col="red", lty=2)
```

```{r}
wisc.hclust.clusters <- cutree()
plot(wisc.hclust.clusters)
```





## Combining methods

```{r}
wisc.pca.hclust <- hclust(dist(wisc.pr$x[,1:7]), method="ward.D2")
```

```{r}
plot(wisc.pca.hclust)
```

```{r}
grps <- cutree(wisc.pca.hclust, k=2)
table(grps)
```


```{r}
table(grps, diagnosis)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps)
```




```{r}
plot(wisc.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16)
```





#Prediction

```{r}
url <- "new_samples.csv"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16)
text(npc[,1], npc[,2], c(1,2), col="white", pch=16)
```

Q17
# The patient in the black zone should be priotitized


//////////////////////////////////////

```{r}
read.csv("WisconsinCancer.csv")
# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv("WisconsinCancer.csv")
```
Examine your input data to ensure column names are set correctly. The id and diagnosis columns will not be used for most of the following steps. Use as.matrix() to convert the other features (i.e. columns) of the data (in columns 3 through 32) to a matrix. Store this in a variable called wisc.data.
```{r}
# Convert the features of the data: wisc.data
as.matrix(wisc.df)
wisc.data <- as.matrix(wisc.df)
```
Assign the row names of wisc.data the values currently contained in the id column of wisc.df. While not strictly required, this will help you keep track of the different observations throughout the modeling process.
```{r}
rownames(wisc.data) <- wisc.df$id
```
Finally, setup a separate new vector called diagnosis to be 1 if a diagnosis is malignant ("M") and 0 otherwise. Note that R coerces TRUE to 1 and FALSE to 0.
```{r}
wisc.df$diagnosis == "M"
as.numeric(wisc.df$diagnosis == "M")
```

```{r}
diagnosis <- as.numeric(wisc.df$diagnosis == "M")
```

EXPLORATORY DATA ANALYSIS
```{r}
#Q1. How many observations are in this dataset?
#shown @ side of data -> 569 obs.
        
#Q2. How many variables/features in the data are suffixed with _mean?
grep("_mean", colnames(wisc.df))
length(grep("_mean", colnames(wisc.df)))
#10

#Q3. How many of the observations have a malignant diagnosis?
sum(as.numeric(wisc.df$diagnosis == "M"))
#212
```

SECTION 2

PERFORMING PCA

It is important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data include:

-The input variables use different units of measurement.
-The input variables have significantly different variances.
Check the mean and standard deviation of the features (i.e. columns) of the wisc.data to determine if the data should be scaled. Use the colMeans() and apply() functions like you’ve done before.
```{r}
colMeans(wisc.data)
```























