---
title: 'Class 8: Machine learning'
author: "Mary Garcia"
date: "April 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##K-means clustering

Let's start with an example of running the **kmeans()** function 

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```


To Do:
- Use the kmeans() function setting k to 2 and nstart=20
- Inspect/print the results

> Q. How many points are in each cluster?
30

```{r}
km$size
km$cluster
km$centers
```

> Q. What ‘component’ of your result object details
 - cluster size? - 30, 30
 - cluster assignment/membership?
 - cluster center?

> Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points

```{r}
#Use $cluster to use cluster assignment
km <- kmeans(x,centers = 2,nstart = 20)
plot(x, col=km$cluster)
points(km$centers, pch = 18, col="blue",  cex = 3)
```









////////////////////////////////////////

## Hierarchial Clustering example

Ex.1
```{r}
# First we need to calculate point (dis)similarity
# as the Euclidean distance between observations
dist_matrix <- dist(x)
# The hclust() function returns a hierarchical
# clustering model
hclust(d = dist_matrix)
# the print method is not so useful here
hc
```
Let's have a closer look...
```{r}
# Our input is a distance matrix from the dist()
# function. Lets make sure we understand it first
dist_matrix <- dist(x)
dim(dist_matrix)
#^NULL
View( as.matrix(dist_matrix) )
dim(x)
#^60 2
dim( as.matrix(dist_matrix) )
#60 60

# Note. symmetrical pairwise distance matrix
```
Interpreting results
```{r}
# Create hierarchical cluster model: hc
hc <- hclust(dist(x))
# We can plot the results as a dendrogram
plot(hc)
# What do you notice?
# Does the dendrogram
# make sense based on
# your knowledge of x?
```

Ex.2
Linking Clusters (Hierarchial clustering)
```{r}
# Using different hierarchical clustering methods
hc.complete <- hclust(d, method="complete")
hc.average <- hclust(d, method="average")
hc.single <- hclust(d, method="single")
plot(hc.complete)
plot(hc.average)
plot(hc.single)
```

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```



////////////////////////////////



We must five the hclust() function a distance matrix not the raw data as input

```{r}
#Distance matrix calculation
d <- dist(x)

# Clustering
hc <- hclust(d)
plot(hc)
```


```{r}
cutree(hc, h=6)
```







```{r}
cutree(hc, k=2)
```





Another example but more real life like with overlapping data

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```




>Q. Use the dist(), hclust(), plot() and cutree()
 functions to return 2 and 3 clusters
>Q. How does this compare to your known 'col' groups?


```{r}
hc <- hclust(dist(x))
plot(hc)
abline(h = 2, col = "red")
abline(h = 2.4, col= "blue")
```

```{r}
gp2 <- cutree(hc, k = 2)
gp3 <- cutree(hc, k = 3)

```

```{r}
gp2
```


```{r}
gp3
```

```{r}
plot(x, col = gp2)
plot(x, col = gp3)
```


```{r}
table(gp2)
```

```{r}
table(gp3)
```

```{r}
table(gp2, gp3)
```


## PCA: Prinical Component Analysis

We will use the **prcomp()** function for PCA
```{r}
#Examle data
## You can also download this file from the class website!
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1)


head(mydata)
```



100 genes in this dataset

```{r}
nrow(mydata)
```

```{r}
ncol(mydata)
```

```{r}
colnames(mydata)
```

Runour PCA analysis on the transpose of our data.

```{r}
prcomp(t(mydata), scale = TRUE)
```

PCA plot

```{r}
## lets do PCA
pca <- prcomp(t(mydata), scale=TRUE)
## A basic PC1 vs PC2 2-D plot
plot(pca$x[,1], pca$x[,2], xlab= "PC1", ylab= "PC2")
```


```{r}
attributes(pca)
```

Calcualte the percent carance captures in each PC

```{r}
## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

head(pca.var.per)
```





```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)"))
```






















```{r}
x <- read.csv("UK_foods.csv")
head(x)
```

```{r}
# Note how the minus indexing works
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
```


```{r}
# Use the prcomp() PCA function 
pca <- prcomp(t(x))
summary(pca)
```

```{r}
# Plot PC1 vs PC2
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
colnames(x)
mycols = c("orange", "red", "blue", "green")
```

```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```









NEW PRACTICE

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```


```{r}
kmeans(x, centers=2, nstart=20)
```


Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points
```{r}
plot
```




